\chapter{Shutdown}


\section{Colophon}

The current version of this document was prepared using \LaTeX{} on Debian. Previous versions of this document were prepared using \LaTeX{} on the following systems:
\begin{itemize}
\item Slackware 13.37, the leetest Linux distro on the planet.
\item Windows 7, using Cygwin's latex packages
\end{itemize}



\printglossary[type=main,title={Glossary},toctitle={Glossary}]





\begin{thebibliography}{1}


\bibitem{event1}
  Eventdriven. \emph{C\&W (CW.) Finals}
  [online] Available at:
  $<$http://is.gd/foF5fC$>$
  [Accessed 28 August 2011].

\bibitem{gannonroita}
  Gannon, Geoff.
  \emph{Free Cash Flow Isn't Everything}
  [online] Available at:
  $<$http://is.gd/nufuGt$>$
  [Accessed 20 May 2012].

\bibitem{bib_fcfy}
  Garcia, Alejandro.
  \emph{How to Calculate  Free Cash Flow Yield}. [online]. Available at:
$<$http://is.gd/4CPCZp$>$
  [Accessed 02 June 2012].

\bibitem{avm}
  Gray, Wesley and Vogel, Jack.
  \emph{Analyzing Valuation Measures: A Performance Horse-Race Over the Past 40 Years}
  Journal of Portfolio Management, Forthcoming.

\bibitem{tlbtbtm}
  Greenblatt, Joel.
  \emph{The Little Book That Beats the Market},
  John Wiley \& Sons, Inc., New Jersey,
  1st Edition,
  2006.

\bibitem{genius}
  Greenblatt, Joel.
  \emph{You Can be a Stock Market Genius},
  Simon \& Schuster Inc., New York,
  1st Edition,
  1997.



\bibitem{mficalc}
  MFI Diary. \emph{The Calculation}.  [online] Available at: 
  $<$http://justadrone.blogspot.com/2011/03/calculation.html$>$
  [Accessed 20 August 2011].

\bibitem{bib:heat}
MFI Diary. \emph{College Funds \& MFI Index Update}.
[online] Available at:
$<$http://is.gd/bM8icZ$>$
[Accessed 21 December 2014].

\bibitem{bib:gtat}
MFI Diary. \emph{GTAT Files for Chapter 11}.
[online] Available at:
$<$http://is.gd/01wjeZ$>$
[Accessed 21 December 2014].


\bibitem{stockcalc}
  Stockopedia. \emph{How Does Magic Formula Investing Work?}.
  [online] Available at:
  $<$http://is.gd/5i4oJX$>$
  [Accessed 20 August 2011].

\bibitem{sce}
  Various.
  \emph{Stock Contest Ends}
  $<$http://is.gd/cMxEdx$>$
  [Accessed 29 May 2012].
  
\bibitem{diet}
  Walker, John. \emph{The Hacker's Diet}.
  [online] Available at:
  $<$http://www.fourmilab.ch/hackdiet/www/hackdiet.html$>$
  [Accessed 28 August 2011].


\end{thebibliography}


\printindex
